Adaptive Video Streaming with Optimized Bitstream Extraction and PID-based Quality Control

### 2015-11-18 tmm.tex
Add citations

\section{Introduction}
\label{sec:intro}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
\IEEEPARstart{V}{ideo} streaming applications are becoming more and more popular on the Internet. Examples of the most common usage scenarios include online video-on-demand (VoD), live broadcasting, video chat, video-based online education, etc. In these applications, video data is constantly delivered to different clients over various transmission channels, and the client media player can begin playing the video before the entire file has been transmitted. Among other features, video streaming makes it easy to watch video anywhere and anytime, which represents the concept of Ubiquitous Multimedia Access (UMA), a very common topic within the multimedia community during the past few years \cite{Gualdi08}. And due to its convenience, video streaming has also attracted a large number of users and become an increasingly important way for people to consume video content \cite{Chen13}.

Although efforts on the new software defined networking (SDN) are made to facilitate efficient video streaming \cite{Egilmez13}, currently traditional best-effort networks still dominate, and the variety of users' network environment remains to be one of the biggest challenges for video streaming. A user might use different types of network (Ethernet, Wi-Fi, Cellular, etc.) and might also switch from one type to another (e.g., from Wi-Fi to cellular network) on going. Even in the same network, the user's bandwidth will change from time to time, influenced by the overall traffic. Within such an environment, bandwidth fluctuation is inevitable. If the video is streamed at a constant bitrate, it would be difficult or impossible for the user to get best watching experience. To cope with this challenge, the video streaming system should be able to automatically adjust the video's bitrate or quality according to the bandwidth change. In other words, the video streaming system needs to be adaptive.

@@

The advantage of simulcast systems like DASH is that they need little modification of existing infrastructure (servers, proxies, caches, firewall compatibility, etc.) and therefore are easy to deploy \cite{Bouten14}. Most commercial video websites, YouTube \cite{YouTube} for instance, use this option to enable quality adaptation. However, the adaptive range and granularity that simulcast provides is limited to the number of streams (YouTube videos have 5 quality levels at most: 240p, 360p, 480p, 720p and 1080p), and the processing and storage of multiple video streams significantly increase time and space cost. SVC streaming systems, on the contrary, do not have above shortcomings. Since an SVC video bitstream can be extracted at the granularity of one NALU (a very small partition of data), it is able to provide seamless scalability and quality adaptation. Besides, performance analysis \cite{SVCPerformance} shows that SVC also has much higher coding efficiency than simulcast, despite some computation and bitrate overhead comparing with single layer stream. Therefore, to adopt SVC in the video streaming framework is a promising option and has attracted much attention in the study \cite{Chuah12}-\cite{Cicalò14}.

@@

\bibitem{Gualdi08}
G. Gualdi, A. Prati, and R. Cucchiara, ``Video Streaming for Mobile Video Surveillance'', in {\em IEEE Transactions on Multimedia}, vol. 10, no. 6, pp.1142-1154, 2008.

\bibitem{Chen13}
Y. Chen, B. Zhang, Y. Liu, and W. Zhu, ``Measurement and Modeling of Video Watching Time in a Large-Scale Internet Video-on-Demand System'', in {\em IEEE Transactions on Multimedia}, vol. 15, no. 8, pp.2087-2098, 2013.

\bibitem{Egilmez13}
H. E. Egilmez, S. Civanlar, and A. M. Tekalp, ``An Optimization Framework for QoS-Enabled Adaptive Video Streaming Over OpenFlow Networks'', in {\em IEEE Transactions on Multimedia}, vol. 15, no. 3, pp.710-715, 2013.

@@

\bibitem{SVCOverview}
H.~Schwarz, D.~Marpe, and T.~Wiegand, ``Overview of the scalable video coding extension of the H.264/AVC standard'', in {\em IEEE Trans. Circuits. Syst. Video Technol.}, vol. 17, no. 9, pp.~1103-1120, Sept. 2007.

\bibitem{Bouten14}
N. Bouten, S. Latré, J. Famaey, W. V. Leekwijck, and F. D. Turck, ``In-Network Quality Optimization for Adaptive Video Streaming Services'', in {\em IEEE Transactions on Multimedia}, vol. 16, no. 8, pp.2281-2293, 2014.

@@

\bibitem{SVCPerformance}
M. Wien , H. Schwarz and T. Oelbaum,  ``Performance Analysis of SVC'',  in {\em IEEE Trans. Circuits Syst. Video Technol.}, vol. 17, no. 9, pp.1194-1203. Sep. 2007.

\bibitem{Chuah12}
S. Chuah, Z. Chen, and Y. Tan, ``Energy-Efficient Resource Allocation and Scheduling for Multicast of Scalable Video Over Wireless Networks'', in {\em IEEE Transactions on Multimedia}, vol. 14, no. 4, pp.1324-1336, 2012.

\bibitem{Zhu13}
Z. Zhu, S. Li, and X. Chen, ``Design QoS-aware multi-path provisioning strategies for efficient cloud-assisted SVC video streaming to heterogeneous clients'', in {\em IEEE Transactions on Multimedia}, vol. 15, no. 4, pp.758-768, 2013.

\bibitem{Dan13}
G. Dan, O. Hadar, R. Ohayon, and N. Amram, ``Live video streaming with adaptive pre-processing by using scalable video coding'', in {\em Proceedings of IEEE International Conference on Consumer Electronics (ICCE)}, pp.588-589, 2013.

\bibitem{Yang14}
E. Yang, Y. Ran, S. Chen, and J. Yang, ``A multicast architecture of SVC streaming over OpenFlow networks'', in {\em Proceedings of IEEE Global Communications Conference (GLOBECOM)}, pp.1323-1328, 2014.

\bibitem{Cicalò14}
S. Cicalò, and V. Tralli, ``Distortion-Fair Cross-Layer Resource Allocation for Scalable Video Transmission in OFDMA Wireless Networks'', in {\em IEEE Transactions on Multimedia}, vol. 16, no. 3, pp.848-863, 2014.

@@

\bibitem{Amonou07}
I.~Amonou, N.~Cammas, S.~Kervadec, S.~Pateux, ``Optimized Rate-distortion Extraction with Quality Layers in the Scalable Extension of H.264/AVC'' in {\em IEEE Trans. Circuits Syst. Video Technol.}, vol.~17, no.~9, pp.~1186--1193, Sep. 2007.

@@

\bibitem{Sun09}
J.~Sun, W.~Gao, D.~Zhao, W.~Li, ``On Rate-distortion Modeling and Extraction of H.264/SVC Fine-Granular Scalable Video'' in {\em IEEE Trans. Circuits Syst. Video Technol.}, vol.~19, no.~3, pp.~323--336, Mar. 2009.

\bibitem{Maani09}
E.~Maani, A.~K. Katsaggelos, ``Optimized Bit Extraction Using Distortion Modeling in the Scalable Extension of H.264/AVC'' in {\em IEEE Trans. Image Processing}, vol.~18, no.~9, pp.~2022--2029, Sep. 2009.

@@

\bibitem{Yang13}
K.~Yang, S.~Wan, Y.~Gong, and Y.~Feng. ``A fast algorithm of bitstream extraction using distortion prediction based on simulated annealing'' In {\em Journal of Visual Communication and Image Representation}, no.7, 752-759, 2013.

\bibitem{Zhang12}
W.~Zhang, J.~Sun, J.~Liu and Z.~Guo, ``Optimized bit extraction of SVC exploiting linear error model,'' {\em Proceedings of IEEE International Symposium on Circuits and Systems}, pp.~1887--1890, Seoul, Korea, May 2012.

@@

\bibitem{HEVC}
G.~J.~Sullivan, J.-R.~Ohm, W.-J.~Han, and T.~Wiegand, ``Overview of the High Efficiency Video Coding (HEVC) standard'' in {\em IEEE Trans. Circuits Syst. Video Technol.}, vol.~22, no.~12, pp.~1648â1667, Dec. 2012.

>>>>

\section{Introduction}
\label{sec:intro}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
\IEEEPARstart{V}{ideo} streaming applications are becoming more and more popular on the Internet. Examples of the most common usage scenarios include online video-on-demand (VoD), live broadcasting, video chat, video-based online education, etc. In these applications, video data is constantly delivered to different clients over various transmission channels, and the client media player can begin playing the video before the entire file has been transmitted. Among other features, video streaming makes it easy to watch video anywhere and anytime, which represents the concept of Ubiquitous Multimedia Access (UMA), a very common topic within the multimedia community during the past few years \cite{Gualdi08}. Due to its convenience, video streaming also attracts a large number of users and becomes an increasingly important way for people to consume video content \cite{Chen13}.

In video streaming, the network condition significantly influences the user's watching experience. To provide better streaming service, some researchers choose to adopt the new software defined networking (SDN) architecture \cite{Egilmez14}. Based on a successful SDN paradigm OpenFlow, noticeable results are achieved, e.g., a solution for improving network throughput in hybrid network \cite{Li14} and a demonstration of adaptive video manycast \cite{Xue15}. Although these studies offer a promising way to optimize network conditions, currently traditional best-effort networks still dominate, and the variety of users' network environment remains to be one of the biggest challenges for video streaming. A user might use different types of network (Ethernet, Wi-Fi, Cellular, etc.) and might also switch from one type to another (e.g., from Wi-Fi to cellular network) on going. Even in the same network, the user's bandwidth will change from time to time, influenced by the overall traffic. Within such an environment, bandwidth fluctuation is inevitable. If the video is streamed at a constant bitrate, it would be difficult or impossible for the user to get best watching experience. To cope with this challenge, the video streaming system should be able to automatically adjust the video's bitrate or quality according to the bandwidth change. In other words, the video streaming system needs to be adaptive.

@@

The advantage of simulcast systems like DASH is that they need little modification of existing infrastructure (servers, proxies, caches, firewall compatibility, etc.) and therefore are easy to deploy \cite{Bouten14}. Most commercial video websites, YouTube \cite{YouTube} for instance, use this option to enable quality adaptation. However, the adaptive range and granularity that simulcast provides is limited to the number of streams (YouTube videos have 5 quality levels at most: 240p, 360p, 480p, 720p and 1080p), and the processing and storage of multiple video streams significantly increase time and space cost. SVC streaming systems, on the contrary, do not have above shortcomings. Since an SVC video bitstream can be extracted at the granularity of one NALU (a very small partition of data), it is able to provide seamless scalability and quality adaptation. Besides, performance analysis \cite{SVCPerformance} shows that SVC also has much higher coding efficiency than simulcast, despite some computation and bitrate overhead comparing with single layer stream. Therefore, to adopt SVC in the video streaming framework is a promising option and has attracted much attention in the study \cite{Chuah12}-\cite{Cicalo14}.

@@

\bibitem{Gualdi08}
G. Gualdi, A. Prati, and R. Cucchiara, ``Video Streaming for Mobile Video Surveillance'', in {\em IEEE Transactions on Multimedia}, vol. 10, no. 6, pp. 1142-1154, 2008.

\bibitem{Chen13}
Y. Chen, B. Zhang, Y. Liu, and W. Zhu, ``Measurement and Modeling of Video Watching Time in a Large-Scale Internet Video-on-Demand System'', in {\em IEEE Transactions on Multimedia}, vol. 15, no. 8, pp. 2087-2098, 2013.

\bibitem{Egilmez14}
H. E. Egilmez and A. M. Tekalp, ``Distributed QoS architectures for multimedia streaming over software defined networks'', in {\em IEEE Transactions on Multimedia}, vol. 16, no. 6, pp. 1597-1609, 2014.

\bibitem{Li14}
S. Li et al., ``Flexible Traffic Engineering (F-TE): When OpenFlow Meets Multi-Protocol IP-Forwarding'', in {\em IEEE Communication Letters}, vol. 18, pp. 1699-1702, Oct. 2014.

\bibitem{Xue15}
N. Xue et al., ``Demonstration of OpenFlow-Controlled Network Orchestration for Adaptive SVC Video Manycast'', in {\em IEEE Transactions on Multimedia}, vol. 17, pp. 1617-1629, Sept. 2015.

@@

\bibitem{SVCOverview}
H.~Schwarz, D.~Marpe, and T.~Wiegand, ``Overview of the scalable video coding extension of the H.264/AVC standard'', in {\em IEEE Trans. Circuits. Syst. Video Technol.}, vol. 17, no. 9, pp. 1103-1120, 2007.

\bibitem{Bouten14}
N. Bouten, S. Latré, J. Famaey, W. V. Leekwijck, and F. D. Turck, ``In-Network Quality Optimization for Adaptive Video Streaming Services'', in {\em IEEE Transactions on Multimedia}, vol. 16, no. 8, pp. 2281-2293, 2014.

@@

\bibitem{SVCPerformance}
M. Wien , H. Schwarz and T. Oelbaum,  ``Performance Analysis of SVC'',  in {\em IEEE Trans. Circuits Syst. Video Technol.}, vol. 17, no. 9, pp. 1194-1203, 2007.

\bibitem{Chuah12}
S. Chuah, Z. Chen, and Y. Tan, ``Energy-Efficient Resource Allocation and Scheduling for Multicast of Scalable Video Over Wireless Networks'', in {\em IEEE Transactions on Multimedia}, vol. 14, no. 4, pp. 1324-1336, 2012.

\bibitem{Zhu13}
Z. Zhu, S. Li, and X. Chen, ``Design QoS-aware multi-path provisioning strategies for efficient cloud-assisted SVC video streaming to heterogeneous clients'', in {\em IEEE Transactions on Multimedia}, vol. 15, no. 4, pp. 758-768, 2013.

\bibitem{Dan13}
G. Dan, O. Hadar, R. Ohayon, and N. Amram, ``Live video streaming with adaptive pre-processing by using scalable video coding'', in {\em Proceedings of IEEE International Conference on Consumer Electronics (ICCE)}, pp. 588-589, 2013.

\bibitem{Yang14}
E. Yang, Y. Ran, S. Chen, and J. Yang, ``A multicast architecture of SVC streaming over OpenFlow networks'', in {\em Proceedings of IEEE Global Communications Conference (GLOBECOM)}, pp. 1323-1328, 2014.

\bibitem{Cicalo14}
S. Cicalo, and V. Tralli, ``Distortion-Fair Cross-Layer Resource Allocation for Scalable Video Transmission in OFDMA Wireless Networks'', in {\em IEEE Transactions on Multimedia}, vol. 16, no. 3, pp. 848-863, 2014.

@@

\bibitem{Amonou07}
I.~Amonou, N.~Cammas, S.~Kervadec, S.~Pateux, ``Optimized Rate-distortion Extraction with Quality Layers in the Scalable Extension of H.264/AVC'' in {\em IEEE Trans. Circuits Syst. Video Technol.}, vol. 17, no. 9, pp. 1186-1193, 2007.

@@

\bibitem{Sun09}
J.~Sun, W.~Gao, D.~Zhao, W.~Li, ``On Rate-distortion Modeling and Extraction of H.264/SVC Fine-Granular Scalable Video'' in {\em IEEE Trans. Circuits Syst. Video Technol.}, vol.~19, no.~3, pp.~323-336, Mar. 2009.

\bibitem{Maani09}
E.~Maani, A.~K. Katsaggelos, ``Optimized Bit Extraction Using Distortion Modeling in the Scalable Extension of H.264/AVC'' in {\em IEEE Trans. Image Processing}, vol.~18, no.~9, pp.~2022-2029, 2009.

@@

\bibitem{Yang13}
K.~Yang, S.~Wan, Y.~Gong, and Y.~Feng. ``A fast algorithm of bitstream extraction using distortion prediction based on simulated annealing'' In {\em Journal of Visual Communication and Image Representation}, no.7, pp. 752-759, 2013.

\bibitem{Zhang12}
W.~Zhang, J.~Sun, J.~Liu and Z.~Guo, ``Optimized bit extraction of SVC exploiting linear error model,'' {\em Proceedings of IEEE International Symposium on Circuits and Systems}, pp.~1887-1890, Seoul, Korea, May 2012.

@@

\bibitem{HEVC}
G.~J.~Sullivan, J.-R.~Ohm, W.-J.~Han, and T.~Wiegand, ``Overview of the High Efficiency Video Coding (HEVC) standard'' in {\em IEEE Trans. Circuits Syst. Video Technol.}, vol.~22, no.~12, pp.~1648-1667, Dec. 2012.
